---
title: "Plots"
author: "Jessica Ernakovich"
date: "1/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) #we changed echo = F globally here so that we didn't have to show our R script
```

```{r, warning = F, message = F, echo = F}
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT) #allows us to make dynamic tables
library(leaflet)
```

# Data Tidying

Escapement data are sources from: Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. doi:10.5063/F1S46Q6M.

```{r, echo = F, results = F}
esc <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", method = "libcurl"), 
                stringsAsFactors = F)

head(esc)
```


Calculate annual, regional total escapement by species.

* `seperate` the date into year - month - day
* `group_by` and `summarize` to calculate the sum over unique region-year-species combinations
* `filter` for salmon species



```{r, echo = F}
annual_esc <- esc %>% 
  separate(sampleDate, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(year = as.numeric(year)) %>% 
  group_by(SASAP.Region, Species, year) %>% 
  summarise(escapement = sum(DailyCount)) %>% 
  filter(Species %in% c("Chinook", "Sockeye", "Pink", "Chum", "Coho")) # %in% filters for any species in the list, so you don't have to do species = "Chinook" | species = "Chum" and so on)

```

```{r, echo = F}
datatable(annual_esc)
```


# Static Plots

```{r, echo = F}
ggplot(data = annual_esc, mapping = aes(x = Species, y = escapement, fill = SASAP.Region)) + 
  geom_col()

# aes function stands for aesthetics, tells gglot which columns to pick and how to map them onto the plot
# most people leave data and mapping arguments implicitly
# for ggplot, you add (+) rather than pipe ( %>% )

# or if you wanted blue for all columns 
#ggplot(data = annual_esc, mapping = aes(x = Species, y = escapement)) + 
  #geom_col(fill = "blue")
```

timeseries of Kodiak escapement by species

```{r, echo = F}
my_theme <- theme_minimal() + 
theme(plot.title = element_text(hjust = 0.5))
```


```{r, echo = F}
ggplot(data = filter(annual_esc, SASAP.Region == "Kodiak"), aes(x = year, y = escapement, 
           color = Species)) + #color sets the outside, or lines, fill sets fill
geom_line() + 
geom_point() + 
ylab("Escapement (no. fish)") + 
xlab("Year") + 
ggtitle("Kodiak Salmon Escapement") + 
my_theme

#can also create your own theme (e.g. theme minimal, defining text sizes, center justification, then call it at the top so you always have the same looking plot)
```


Plot timeseries of escapement by species for all regions

```{r, echo = F, fig.height = 10, fig.width = 6}
esc_plot <- ggplot(annual_esc, aes(x = year, y = escapement, color = Species)) +
  geom_line() +
  geom_point() +
  facet_wrap(~SASAP.Region, scales = "free", ncol = 2) + # ~ is a formula in R that defines the independent ~ dependent variables
  #free allowed the axes to not be consistent across all the graphs (can also have free_y,etc)
  #facet_wrap allows you to make the same type of graph for many different panels of data... the same plot
  # gridExtra function grid.arrange allows you to set up a multipanel plot, which is especially helpful if you need to make plots that are different
  my_theme

esc_plot

ggsave("figures/region_escapement.png", 
       plot = esc_plot, 
       height = 10, 
       width = 8, 
       units = "in")

```

to write a datafile

```{r, echo = F}
write.csv(annual_esc, "derived_data/annual_escapement.csv", 
          row.names = F)

#if you don't want to push the csv to your GitHub respository, you can add this to your .gitignore file
```




# Interactive Map

Make a map using leaflet  

Find unique locations with lat/long values

```{r}
locations <- esc %>%
  distinct(Location, Latitude, Longitude) %>%  #distinct will pull out unique combinations of whatever values you give it
drop_na() # will drop a row if ANY of the values are missing (even the  uniqueID, here location)

datatable(locations)


```

```{r}
leaflet(locations) %>%  #leaflet uses the pipe (unlike ggplot)
  addTiles() %>% #adds the background map
  addMarkers(~ Longitude, ~ Latitude, popup = ~ Location)

# you can even put ggplot plots into leaflet as popups
```






